EuPPollNet: A European database of plant-pollinator networks

### 2024-08-22 Manuscript/Figures_and_Tables/Exact_location.qmd
adressing comments from colleagues

---
title: "EuPPollNet: A European database of plant-pollinator networks"
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}
        \usepackage{lineno}
        \linenumbers
bibliography: references.bib
csl: apa_6th.csl
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(sf) #For handling coordinates and plotting
library(scales) #For plotting (decimals on axes)
library(lubridate) #To operate with dates
library(tidyr) #Reshape data (wide format)
library(stringr) #Fix/modify strings
library(forcats)
library(readr)
library(phytools)
#Load libraries for plotting
library(giscoR) #Maps
library(ggstar) #cool shapes as points
library(ggplot2) #general plotting
library(ggtree) #phylo trees
library(ggpattern) #cool shapes in barplots
library(patchwork) #plot organisation
library(cowplot) #plot organisation
library(viridis) #colours
library(ggnewscale)
library(ggtreeExtra)
library(kableExtra)

#In case of conflicts select the main dplyr functions
library(conflicted)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)

```

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}

#Read data
data = readRDS("../Data/Working_files/Interactions_uncounted.rds")%>% 
mutate(Network_id = paste0(Study_id, Network_id))

#Bee-syrphid Masterlist
master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
#Load butterfly masterlist
master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
                    header = T)
#Prepare table with data structure (as it would be made available to others)
safenet_database = readRDS("../Data/3_Final_data/Interaction_data.rds")

#Work around to get the three cols
#Note that month 1 and day 1 indicates the presence of a missing value
#At least in Europe :)
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
                Month = lubridate::month(Date), 
                Day = lubridate::day(Date))

#Country number
country_number = data %>% 
select(Country) %>% 
n_distinct()
#Study number
study_number = data %>% 
select(Study_id) %>% 
n_distinct()
#Average study number
studies_by_country = data %>% 
group_by(Country) %>% 
summarise(study_number = n_distinct(Study_id)) %>% 
summarise(mean(study_number)) %>% 
pull()
#$\bar{x}_\frac{studies}{country}$ = `r studies_by_country`
#Sampling method
Sampling_method = data %>% 
group_by(Study_id) %>% 
distinct(Sampling_method) %>%  
ungroup() %>% 
count(Sampling_method) %>% 
mutate(percent_method = n/sum(n) *100) %>% 
mutate(Sampling_method = recode_factor(Sampling_method, "Transect" = "transects"))
#1st most common sampling methodology
most_common_method = Sampling_method %>% 
filter(percent_method == max(percent_method)) 
#Minimum year
min_year = data %>% 
select(Year) %>% 
slice(which.min(Year)) %>% 
pull()
#Maximum year
max_year = data %>% 
select(Year) %>% 
slice(which.max(Year)) %>% 
pull()
#Number of interactions
interaction_number = nrow(data)
#Check number of accepted pollinators
pollinator_spp_number = data %>% 
select(Pollinator_rank, Pollinator_accepted_name)%>%
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name) %>% 
n_distinct()
#Check number of plant species
plant_spp_number = data %>% 
select(Plant_rank, Plant_status, 
Plant_matchtype, Plant_accepted_name) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name) %>% 
n_distinct()
#Extract percentage of studies conducted during a single year
dates = data %>% 
select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
mutate(Date = dmy(paste(Day, Month, Year, sep = "-"))) %>% 
group_by(Study_id) %>% 
summarise(Count = n_distinct(Year))
#Filter studies with only one year
one_year = dates %>% 
filter(Count == 1) %>% 
pull() %>% 
length() / nrow(dates) * 100
#Obtain date information
dates = data %>% 
select(Study_id, Network_id,Sampling_method, Day, Month, Year) %>% 
mutate(Date = dmy(paste(Day, Month, Year, sep = "-")))

dates1 = dates %>% 
group_by(Study_id, Network_id, Sampling_method) %>% 
summarise(Min_date = min(Date), 
          Max_date= max(Date), 
          Sampling_days = n_distinct(Date)) %>% 
mutate(Sampling_period = as.numeric(Max_date - Min_date)) %>% 
mutate(Study_number = as.integer(str_extract(Study_id, "[^_]+")), 
       .before=Study_id) %>% 
arrange(Study_number) %>% 
mutate(Sampling_period = case_when(!is.na(Sampling_period) ~ Sampling_period + 1,
TRUE ~ NA)) %>% 
mutate(Sampling_days = case_when(!is.na(Sampling_period) ~ Sampling_days,
TRUE ~ NA))
#Extract min and max number of sampling days
min_sampling_day = dates1 %>% pull(Sampling_days) %>% min(na.rm = TRUE)
max_sampling_day = dates1 %>% pull(Sampling_days) %>% max(na.rm = TRUE)
#Extract average of sampling days across networks and studies
average_sampling_days = dates1 %>% 
ungroup() %>% 
group_by(Study_id) %>% 
summarise(mean_sampling_days = mean(Sampling_days, na.rm = TRUE)) %>% 
ungroup() %>% 
summarise(mean(mean_sampling_days, na.rm = TRUE)) %>% 
pull()

#Check number of distinct networks
n_networks = data %>% 
mutate(Network_id = paste0(Network_id,"_", Year)) %>% 
distinct(Network_id) %>% 
nrow(.)

#Check number of distinct sampling sites
n_sites = data %>% 
select(Latitude, Longitude) %>% 
distinct() %>% 
nrow(.)
#Check number of distinct bioregions
n_bioregion = data %>% 
select(Bioregion) %>% 
distinct()  %>% 
nrow(.)
#Read author list 
authors = readr::read_csv("../Data/Manuscript_info/Authorship_ordered.csv")
#Print authors
authors = authors %>% 
summarise(col = paste(Coauthor_name_aff, collapse="| ")) %>% 
pull()
#Read affiliation list 
aff_list = readr::read_csv("../Data/Manuscript_info/Affiliation_list.csv")
aff_list = aff_list %>% 
summarise(col = paste0(Aff_number_Aff, "\\",sep="")) %>% 
pull()

#Interactions at species level
species_level = data %>% 
filter(Plant_rank == "SPECIES") %>% 
filter(Pollinator_rank == "SPECIES") %>% 
filter(Plant_status == "ACCEPTED") %>% 
filter(Pollinator_status == "ACCEPTED") %>% 
nrow(.)
species_level_int = species_level / nrow(data) * 100
#Read here plant and pollinator coverage
bee_coverage = readRDS("../Data/Manuscript_info/bee_coverage.RData")
lepidoptera_coverage = readRDS("../Data/Manuscript_info/lepidoptera_coverage.RData")
syrphid_coverage = readRDS("../Data/Manuscript_info/syrphid_coverage.RData")
plant_coverage = readRDS("../Data/Manuscript_info/plant_coverage.RData")

#Calculate mean coverage and deviation
coverage = c(bee_coverage, syrphid_coverage, lepidoptera_coverage)
mean_coverage = mean(coverage)
sd_coverage = sd(coverage)

#Main orders percentage
o = c("Hymenoptera", "Lepidoptera", "Diptera", "Coleoptera")
main_orders = data %>% 
filter(Pollinator_order %in% o)
main_orders_percent = nrow(main_orders) / nrow(data) * 100

#Prepare table with all info 
#Create variables
vars = colnames(safenet_database)
#Create table
d = tibble(Variable = vars)

#Add descriptor
my_tibble = d %>%
  mutate(Description = case_when(
    Variable == "Study_id" ~ "Identifier of the study",
    Variable == "Network_id" ~ "Identifier of a site sampled within a study",
    Variable == "Sampling_method" ~ "Type of plant-pollinator sampling",
    Variable == "Authors_habitat" ~ "Type of habitat as described by the authors",
    Variable == "EuPPollNet_habitat" ~ "Type of habitat homogenized across studies",
    Variable == "Bioregion" ~ "European biogiographical regions",
    Variable == "Country" ~ "Country where the plant-pollinator interaction was observed",
    Variable == "Locality" ~ "Locality where the plant-pollinator interaction was observed",
    Variable == "Latitude" ~ "North-south position of the observed interaction in decimal degrees",
        Variable == "Longitude" ~ "East-west position of the observed interaction in decimal degrees",
    Variable == "Date" ~ "Year, month and day when the observation took place",
    Variable == "Interaction" ~ "Number of interactions. By default is 1 as interactions are provided ungrouped",
    Variable == "Plant_original_name" ~ "Plant species name given by the authors",
    Variable == "Plant_accepted_name" ~ "Harmonized plant species name in the database",
    Variable == "Plant_rank" ~ "Taxonomic rank of the observation",
    Variable == "Plant_order" ~ "Order taxonomic rank of the observed plant species",
    Variable == "Plant_family" ~ "Family taxonomic rank of the observed plant species",
    Variable == "Plant_genus" ~ "Genus taxonomic rank of the observed plant species",
    Variable == "Plant_unsure_id" ~ "Category to indicate if the plant species name is unsure (Yes) or not (No)",
    Variable == "Plant_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
    Variable == "Pollinator_original_name" ~ "Pollinator species name given by the authors",
    Variable == "Pollinator_accepted_name" ~ "Harmonized pollinator species name in the database",
    Variable == "Pollinator_rank" ~ "Taxonomic rank of the observation",
    Variable == "Pollinator_order" ~ "Order taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_family" ~ "Family taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_genus" ~ "Genus taxonomic rank of the observed pollinator species",
    Variable == "Pollinator_unsure_id" ~ "Category to indicate if the pollinator species name is unsure (Yes) or not (No)",
    Variable == "Pollinator_uncertainty_type" ~ "If the name is unsure, type of species uncertainty is provided",
        Variable == "Flower_data" ~ "Floral data availability (Yes) or (No)",
        Variable == "Flower_data_merger" ~ "Column to merge floral data with the interaction dataset",

    TRUE ~ "Other"
  ))

#Check number of species and percentage of interactions of poll. groups
main_orders = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera")

order_info = data %>% 
select(Pollinator_order, Pollinator_accepted_name) %>% 
filter(Pollinator_order %in% main_orders) %>% 
group_by(Pollinator_order) %>% 
summarise(Interactions = length(Pollinator_accepted_name)/ nrow(.) *100,
          Species = n_distinct(Pollinator_accepted_name))

Hymenoptera_interactions = order_info %>% filter(Pollinator_order == "Hymenoptera") %>%
select(Interactions) %>% pull()

#Check percentage of studies with flower counts
floral_counts_percent = data  %>%
select(Study_id, Flower_data) %>% 
distinct() %>% 
group_by(Flower_data) %>% 
summarise(Percentage = n()/ nrow(.)*100) %>% 
filter(Flower_data == "Yes")

#Check percentage of apis mellifera records
all_records = data %>% nrow(.) 
apis_records = data %>% 
filter(Pollinator_accepted_name == "Apis mellifera") %>% 
nrow(.)
apis_percent = apis_records/all_records * 100
#Check average pecentage of interactions of apis by net
a = data %>% 
group_by(Network_id) %>% 
summarise(n_interactions = length(Pollinator_accepted_name))
b = data %>% 
group_by(Network_id) %>% 
filter(Pollinator_accepted_name == "Apis mellifera") %>% 
summarise(n_interactions_apis = length(Pollinator_accepted_name))
c = left_join(a,b)
apis_average_network = c %>% 
mutate(apis_percent_net = n_interactions_apis/ n_interactions *100) %>% 
ungroup() %>% 
summarise(mean_apis_percent_net = mean(apis_percent_net, na.rm=T)) %>% 
pull()

#Check number of species from Hymenop, Dipte, Lepidop and Coleop

pollinator_spp = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
distinct(Pollinator_accepted_name) %>% 
nrow(.)

pollinator_spp_main_orders = data %>% 
select(Pollinator_order, Pollinator_rank, Pollinator_accepted_name, Interaction) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Hymenoptera"|
       Pollinator_order == "Diptera"|
       Pollinator_order == "Lepidoptera"|
       Pollinator_order == "HymenoColeopteraptera") %>% 
distinct(Pollinator_accepted_name) %>% 
nrow(.)

spp_percent_main_order = pollinator_spp_main_orders/pollinator_spp *100

#Calculate percentage of studies that recorded the main 4 orders
metadata = readRDS("../Data/3_Final_data/Metadata.rds")
#Some studies are divided by year
#Get an average of the pollinator percentages across years first
#Fix NA's
metadata_study = metadata %>%
group_by(Study_id) %>% 
mutate(Hymenoptera_percent = as.numeric(na_if(Hymenoptera_percent, "NA"))) %>%  
mutate(Diptera_percent = as.numeric(na_if(Diptera_percent, "NA"))) %>%  
mutate(Lepidoptera_percent = as.numeric(na_if(Lepidoptera_percent, "NA"))) %>%  
mutate(Coleoptera_percent = as.numeric(na_if(Coleoptera_percent, "NA"))) 
#Calculate average
mean_study = metadata_study  %>% 
summarise(Hymenoptera_percent = mean(Hymenoptera_percent, na.rm = T),
          Diptera_percent = mean(Diptera_percent, na.rm = T),
          Lepidoptera_percent = mean(Lepidoptera_percent, na.rm = T),
          Coleoptera_percent = mean(Coleoptera_percent, na.rm = T)) %>% 
 mutate(across(c(Hymenoptera_percent, Diptera_percent, Lepidoptera_percent, Coleoptera_percent),
                ~ ifelse(is.nan(.), NA, .))) 
#Now calculate percentage of studies that record any of these 4 orders
mean_percent_order = mean_study %>%
  summarise(
    Hymenoptera_percent_metaweb = sum(!is.na(Hymenoptera_percent)) / nrow(.) * 100,
    Diptera_percent_metaweb = sum(!is.na(Diptera_percent)) / nrow(.) * 100,
    Lepidoptera_percent_metaweb = sum(!is.na(Lepidoptera_percent)) / nrow(.) * 100,
    Coleoptera_percent_metaweb = sum(!is.na(Coleoptera_percent)) / nrow(.) * 100)

Hymenoptera_percent_metaweb = mean_percent_order$Hymenoptera_percent_metaweb
Diptera_percent_metaweb = mean_percent_order$Diptera_percent_metaweb
Lepidoptera_percent_metaweb = mean_percent_order$Lepidoptera_percent_metaweb
Coleoptera_percent_metaweb = mean_percent_order$Coleoptera_percent_metaweb

#Check percentages of hymenoptea groups
hymenoptera_percent_by_group = metadata %>% 
ungroup() %>% 
distinct(Study_id, Taxa_recorded) %>% 
summarise(Bee_studies = sum(str_detect(Taxa_recorded, "Bees")/ nrow(.) * 100),
          Bumblebee_studies = sum(str_detect(Taxa_recorded, "Bumblebees")/ nrow(.) * 100),
          Hymenoptera_studies = sum(str_detect(Taxa_recorded, "Hymenoptera")/ nrow(.) * 100))

bee_percent = hymenoptera_percent_by_group$Bee_studies
bumblee_percent = hymenoptera_percent_by_group$Bumblebee_studies
all_hymnoptera_percent = hymenoptera_percent_by_group$Hymenoptera_studies

#Check percentages of diptera groups
diptera_percent_by_group = metadata %>% 
ungroup() %>% 
distinct(Study_id, Taxa_recorded) %>% 
summarise(Syrphid_studies = sum(str_detect(Taxa_recorded, "syrphids")/ nrow(.) * 100),
          Diptera_studies = sum(str_detect(Taxa_recorded, "Diptera")/ nrow(.) * 100),
          other_diptera_studies = sum(str_detect(Taxa_recorded, "syrphids, bombylids") |
                                       str_detect(Taxa_recorded, "syrphids, conopids, tachinid flies"))/ nrow(.) * 100)

syrphid_percent = diptera_percent_by_group$Syrphid_studies
all_diptera_percent = diptera_percent_by_group$Diptera_studies
other_diptera_percent = diptera_percent_by_group$other_diptera_studies

```

\vspace*{-14mm}

`r authors`

**Corresponding author**= barragansljose@gmail.com\
*All authors excluding the first four and last are ordered alphabetically*

\begingroup
\small

`r aff_list`

\small
\endgroup

\pagebreak

# Abstract
\vspace*{-5mm}

**Motivation:** Pollinators play a crucial role in maintaining Earth’s terrestrial biodiversity and human food production by mediating sexual reproduction for most flowering plants. Indeed, the network of interactions formed by plants and pollinators constitutes the backbone of plant-pollinator community stability and functioning. However, rapid human-induced environmental changes are compromising the long-term persistence of plant-pollinator interaction networks. One of the major challenges for pollinator conservation is the lack of robust generalisable data capturing how plant-pollinator communities are structured across space and time. Here, we present the EuPPollNet (European Plant-Pollinator Networks) database, a fully open and reproducible European-level database containing harmonized taxonomic data on plant-pollinator interactions referenced in both space and time, along with other ecological variables of interest. This database offers an open workflow that allows researchers to track data-curation decisions and edit them according to their preferences. We present the taxonomic and sampling coverage of EuPPollNet, and summarize key structural properties in plant-pollinator networks. We hope EuPPollNet will stimulate future research that fills the taxonomic, ecological, and geographical data gaps on plant-pollinator interactions that we have identified. Further, the variation in the structure of the networks in EuPPollNet provides a strong basis for future studies aimed at quantifying drivers of plant-pollinator network change and guiding future conservation planning for plants and pollinators.

**Main Types of Variables Included:** EuPPollNet contains `r prettyNum(interaction_number, big.mark=",")` interactions between plants and pollinators from `r prettyNum(n_networks, big.mark=",")` distinct networks (i.e., distinct sampling event in space or time), which belong to `r study_number` different studies distributed across `r country_number` European countries. In addition, information about sampling methodology, habitat type, bio-climatic region, and further taxonomic rank information for both plant and pollinator species are also provided (i.e., genus, family and order).

**Spatial location and grain:** The database contains `r prettyNum(n_sites, big.mark=",")` different sampling locations from natural and anthropogenic habitats that fall in `r n_bioregion` different bio-climatic regions. All records are geo-referenced and presented in the World Geodetic System 1984 (WGS84).

**Time period and grain:** Species interaction data was collected between `r min_year` and `r max_year`. All records are time-referenced and most of the studies documented interactions within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%).

**Major taxa and level of measurement:** The database contains interaction data at the species level for `r format(round(species_level_int, 2), nsmall = 2)`% of the records, including a total of `r prettyNum(plant_spp_number, big.mark=",")` plant and `r prettyNum(pollinator_spp_number, big.mark=",")` pollinator species. The database covers `r format(round(plant_coverage, 2), nsmall = 2)`% of the European species of flowering plants, `r format(round(bee_coverage, 2), nsmall = 2)`% of bees, `r format(round(lepidoptera_coverage, 2), nsmall = 2)`% of butterflies, and `r format(round(syrphid_coverage, 2), nsmall = 2)`% of syrphid species at the European level.

**Software format:** The database was built with the R programming language and is stored as “.rds” and “.csv” formats. The construction of the database is fully reproducible and can be accessed at the following link: <https://github.com/JoseBSL/EuPPollNet.git>.

**KEYWORDS**\
\begingroup
\small
Angiosperms, connectance, flowering plants, nestedness, plant-pollinator networks, pollinators, species interactions 
\endgroup
\pagebreak

# 1 | INTRODUCTION

Plant-pollinator interactions involve a great diversity of species, largely attributed to their coevolutionary history [@ollerton2017], and are critically important for terrestrial biodiversity and economic productivity. The synergistic effects of climate change with other global change pressures are threatening worldwide biodiversity [@sala2000; @bellard2014], including plant and pollinator species as well as their interactions [@goulson2015; @settele2016; @eichenberg2021]. Under this scenario, the increasing availability of biodiversity data plays a major role in our ecological understanding of species status, trends, and conservation [@heberling2021; @zattara2021]. However, our knowledge of plant and pollinator species and their network of interactions still exhibits major temporal, spatial and taxonomic biases [@archer2014; @troia2016; @poisot2021; @marshall2024], limiting our ability to effectively protect their biodiversity.

The interactions between different plant and pollinator species within a community form complex networks. Macro-ecological analyses of the topology of these networks have revealed common properties, such as truncated power-law degree distributions [@jordano2003] or nestedness [@bascompte2003]. Large-scale analyses across multiple studies can quantify patterns across geographic regions [@olesen2002; @traveset2016] or environmental gradients [@ramos2010; @rech2016; @saunders2023] that cannot be examined in a single study. Although macro-ecological approaches that use ecological interactions make significant contributions to knowledge [@windsor2023], such synthesis work must consider variation across studies in the spatio-temporal nature of the data [@burkle2011; @garcia2024]. For instance, plant-pollinator studies tend to differ in sampling effort and methodology which affect the structure of the resulting plant-pollinator networks [@gibson2011; @jordano2016; @schwarz2020]. Most plant-pollinator networks have unobserved interactions [@olesen2011; @chacoff2012], and thus research that attempts to synthesize across published studies must have access to raw data on interactions in order to statistically account for sampling effort and completeness. This emphasizes the importance of providing data in its rawest possible form in datasets that will be utilized for synthesis and macro-ecological studies.

Europe is one of the continents with a larger amount of available biodiversity data [@proencca2017], yet still exhibits major gaps [@wetzel2018; @bennett2018]. While species checklists need to be treated carefully, especially at a macro-ecological scale [@grenie2023], the growing number of European plant and pollinator checklists [@reverte2023], along with occurrence data [@zattara2021], is setting a foundation for the conservation of its flora and their pollinators. However, species richness is just one component of biodiversity and documenting the interaction between plants and pollinators is essential for understanding biodiversity change [@jordano2016]. Numerous works have studied plant-pollinator interactions in the last decades, generating thousands of plant-pollinator interaction networks worldwide. Several initiatives have tried to integrate plant-pollinator interaction data into databases such as *Mangal* [@poisot2016] or *GloBI* [@poelen2014], resulting in numerous large scale comparative analyses that have enhanced our understanding of the ecology of plants and pollinators [e.g., European wild bee data trends\; @marshall2024]. Despite all these resources, Europe lacks accessible harmonized plant-pollinator interaction data that allow researchers to evaluate plant-pollinator interactions at a European level. For example, only over a dozen of European plant-pollinator networks are included in *Mangal*, while *GloBI* focuses on pairwise interactions disconnected from the community context. Assembling and curating the existing information on EU plant-pollinator networks will guide research efforts, conservation planning and will set a foundation for future global change research.

Here, we present the European Plant-Pollinator Networks database (EuPPollNet), which contains harmonized information on plant-pollinator interactions at the European level. The pollinator taxonomic groups include the main orders of entomofauna that visit and pollinate flowering plants in Europe. These comprise insect species from the orders Hymenoptera, Diptera, Lepidoptera, and Coleoptera, accounting for almost the totality of recorded interactions in EuPPollNet (`r format(round(main_orders_percent, 2), nsmall = 2)`%). To understand the scope of the database, we examined the taxonomic and sampling coverage of the different plant and pollinator species at the European level with the help of the most up-to-date species checklists and rarefaction analyses. In addition, for bees and plants, we evaluated if there is a phylogenetic signal in the presence-absence of interaction data.

EuPPollNet contains one of the largest sets of plant-pollinator networks collated to date, providing a unique opportunity to examine the prevalence of key structural metrics across networks. For example, despite the large theoretical literature on the meaning of a nested structure in plant-pollinator networks [@bascompte2007; @guimaraes2020], where specialists species interact only with subsets of generalists species, this pattern has only been empirically evaluated with a relatively small number of networks [@bascompte2003; @staniczenko2013; @payrato2019], and is still debated how structural metrics such as connectance and nestedness change across latitudes and bioclimatic regions [@olesen2002; @trojelsgaard2013; @song2017].

Overall, EuPPollNet aims to cover a wide range of taxonomic groups and habitats, while also providing other variables of interest that define the ecological context and sampling methods of the study. In addition, EuPPollNet offers a transparent and accessible workflow of its data management and species harmonization that allows the database to be reused and to expand over time. This database provides a large number of community-level networks with curated and harmonized data, distinguishing it from other currently available resources that contain plant-pollinator interactions. We expect that EuPPollNet can be used to evaluate macro-ecological processes in plant-pollinator networks, guide conservation planning, and set a baseline for global change research.

# 2 | METHODS

## Data acquisition

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were "*plant-pollinator interactions*" and "*plant-pollinator networks*". To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting them as the contact between a pollinator and the reproductive structure of a specific plant being sampled (i.e., phyto-centric networks).

## Dataset description

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders `r format(round(100-spp_percent_main_order, 2), nsmall = 2)`% are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="105%", fig.cap= "\\textbf{Figure 1}. (\\textbf{a}) Locations of the studies in EuPPollNet across the European continent, showing the total number of pollinator (i.e., orange heptagon) and plant (i.e., green circles) species per study. The sizes of these shapes are proportional to the respective species counts. For visualization purposes, we have selected only a single location per study. (\\textbf{b}) Number of studies by year in EuPPollNet. (\\textbf{c}) Proportion of species and interactions across the four main pollinator orders in EuPPollNet, excluding interactions from \\textit{Apis mellifera}. The total number of species and interactions is indicated in parentheses at the bottom."}

@@

#Prepare coordinates
for_plotting = data %>% 
select(Study_id, Latitude, Longitude) %>% 
distinct()
#Get countries and transfrom coordinates
countries <- gisco_get_countries(
resolution = 20) %>%
st_transform(3035) %>% 
sf::st_as_sf(crs = 4326)

#Number of plant and poll species per network
#-----------------------#
#Pollinators------
#-----------------------#
#Check unique number of pollinator species
PollNumber_by_ID = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Study_id) %>% 
filter(Pollinator_rank == "SPECIES") %>% 
select(Pollinator_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Pollinator_counts = n_distinct(Pollinator_accepted_name))
#Merge now with spp counts by network dataset
poll_data = left_join(PollNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
poll_data$coords <- data.frame(
Longitude = poll_data$Longitude, Latitude = poll_data$Latitude)
#Set coordinates
poll_data$coords_sf = st_as_sf(poll_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
poll_data1 =  poll_data %>% 
distinct(Study_id, .keep_all = TRUE)

#-----------------------#
#Plants-----
#-----------------------#
#Check unique number of plant species
PlantNumber_by_ID = data %>% 
select(Plant_rank, Plant_accepted_name, Study_id) %>% 
filter(Plant_rank == "SPECIES") %>% 
select(Plant_accepted_name, Study_id) %>% 
group_by(Study_id) %>% 
summarise(Plant_counts = n_distinct(Plant_accepted_name))
#Merge now with spp counts by network dataset
plant_data = left_join(PlantNumber_by_ID, for_plotting, by = "Study_id")
#Create duplicate cord columns to operate on them
plant_data$coords <- data.frame(
Longitude = plant_data$Longitude, Latitude = plant_data$Latitude)
#Set coordinates
plant_data$coords_sf = st_as_sf(plant_data$coords, 
  coords = c("Longitude", "Latitude"), crs = 4326)
#Plot map of Europe
#Very noisy with many coordinates per point,
#Let's select a unique coordinate by study
plant_data1 =  plant_data %>% 
distinct(Study_id, .keep_all = TRUE)

#Add vars
plant_data1$Group = "Plant"
poll_data1$Group = "Pollinator"
#Rename cols
plant_data2 = plant_data1 %>% 
rename(Counts = Plant_counts)
poll_data2 = poll_data1 %>% 
rename(Counts = Pollinator_counts)
#Bind rows
all_data = bind_rows(plant_data2, poll_data2)

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_star(data = all_data, alpha=1,aes(x= Longitude, y=Latitude,
           size=Counts, starshape = Group, color= Group))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
      legend.position = ("top"),
      legend.box = "vertical",
      legend.background = element_blank(),
      legend.box.background = element_blank(),
      legend.key = element_blank(), 
      legend.title = element_text(face ="bold",  size = 6),
      legend.box.spacing = unit(0.2, "cm"),
      legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(-10,-10,-10,-10),
      axis.text = element_text(size=6,                  face="bold"),
      axis.title = element_text(face="bold", size = 8),
         legend.text = element_text(size = 6),                           legend.direction="horizontal",
         legend.spacing.y = unit(-0.3, "cm"),
      legend.title.align=0.5) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(name= "Species counts",range = c(0, 3.75))+
scale_color_manual(values=c("#009E73", "#D55E00"),
         guide="none") +
scale_starshape_manual(name = "Taxonomic group",             values = c(15, 7),
             guide=guide_legend(reverse=TRUE,
             override.aes=list(
                 size=2,
                 color=c("Pollinator"= "#D55E00","Plant"="#009E73")))) +
  ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(color = "none", size=guide_legend(nrow=1,override.aes = list(starshape =c(15,15, 15), size=c(0.8,1.8,2.8)))) + 
ggtitle("(a)") +
theme(plot.title = element_text(vjust = -5, hjust=0.057, size=10))

>>>>

---
latex-tinytex: false
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(ggplot2) #For plotting
library(giscoR) #For plotting
library(sf) #For handling coordinates and plotting
library(patchwork)#For plotting (binding plots)
library(ggstar) #For plotting (cool shapes)
library(scales) #For plotting (decimals on axes)
library(tidyr)
library(viridis)
```

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}

#Load data
data = readRDS("../../Data/3_Final_data/Interaction_data.rds")

@@
